---
output:
  pdf_document: default
  html_document: default
---
# Reproducible Research - Activity Monitoring (Project 1 Resubmission)
 
================================================================================

There is a rise in the so-called "quantified self" movement, where people
track their physical activity and collect large data sets in order to find
patterns about themselves in pursuit of self-improvement.
 
However, there are often large sets of raw data being collected that are not
processed, and as a result often go under-utilized.
 
In this project, data from an anonymous individual were collected over a
two-month time period in October and November 2012 using a personal activity
monitoring device.
 
## - Loading and Transforming Activity Monitoring Data
 
```{r}
echo = TRUE
# Load ggplot2 functions
 
library(ggplot2)
 
# When the .csv file is in your local R/RStudio directory and already unzipped:
 
ActMon <- read.csv("/Users/alex/Desktop/Coursera/Reproducible_Research/activity.csv")
```
 
These data contain the following three variables:
 
- steps (5-minute interval inclusive of the number of steps)
- date (YYYY-MM-DD format for when data were collected)
- interval (identifies the particular time interval data were collected)
 
The following questions have been addressed in this markdown document:
 
## - What is the mean total number of steps taken per day?
 
```{r}
echo = TRUE
# Find the total number of steps taken per day, then summary to find the 
# Mean and Median at the same time, along with other distribution data
 
T_Steps <- aggregate(steps ~ date, ActMon, sum)
print(summary(T_Steps))
```
 
## - How does the distribution of values for total number of steps change
##   over the course of the period in quesion?
```{r}
echo = TRUE
# Simple histogram demonstrating the number of steps was close to Mean/Median
 
hist(T_Steps$steps, breaks = 16, xlab = "Total Steps per Day", main = "Frequency
      of Daily Step Totals from October-November 2012")
```
 
## 3. What is the average daily activity pattern for this individual like?
 
```{r}
echo = TRUE
# Compute the mean of all steps based on the interval in which they occurred,
# then save the output as an object
Int_Act <- aggregate(steps ~ interval, ActMon, mean)
 
# Plot the average activity pattern seen in a 24-hour period of time
ggplot(data = Int_Act) +
  geom_line(aes(interval,steps)) +
  xlab("5-Minute Interval over 24 hour Period") +
  ylab("Number of Steps") +
  ggtitle("Average Step Count by Time of Day (in 5 min Increments)")
 
```
## - At what 5-interval point is the highest average number of steps taken?
```{r}
echo = TRUE
# Display the interval where most number of average steps are taken
 
print(Int_Act[which.max(Int_Act$steps),])
```
We can see in the 835 interval that the highest average number of steps is
about 206.
 
## - How many entries are missing data?
```{r}
echo = TRUE
# Determining the total missing values by the total number flagged by the
# system
 
print(colSums(is.na(ActMon)))
```
 
There are 2,304 steps entries that are NA.
 
## - What do the data look like when missing values are changed to the original
##   data's mean or median values?
```{r}
echo = TRUE
# Start transforming NA values in steps into the mean of the original data set
# for values in the original data set where NA values are present
 
ActMon -> ActMon2
 
ActMon2$steps <- ifelse(is.na(ActMon2$steps) == TRUE,
                      Int_Act$steps[Int_Act$interval %in%
                                      ActMon2$interval],
                      ActMon2$steps)
print(head(ActMon2))
```
## - How did the Mean and Median change from the initial data?
 
```{r}
echo = TRUE
# Find the total number of steps taken per day, then summary to find the 
# Mean and Median at the same time, along with other distribution data
 
T_Steps2 <- aggregate(steps ~ date, ActMon2, sum)
print(summary(T_Steps2))
```
 
The steps distribution data are nearly similar, except the Median here
is 1 higher than the original data.
 
## - How does the distribution of values for total number of steps change
##   over the course of the period in quesion? 
```{r}
echo = TRUE
# Simple histogram demonstrating the number of steps was close to Mean/Median
# using the transformed data set T_Steps2
 
hist(T_Steps2$steps, breaks = 16, xlab = "Total Steps per Day", main = "Frequency
    of Daily Step Totals (Transformed) from October-November 2012")
```
Based on this new plot, adding the mean values in place of NA values biased 
this histogram to have a higher frequency of values close to the mean.
 
##  Are there any activity/ or differences between weekdays and holidays?
 
```{r}
echo = TRUE
# Create a new column in the original data set ActMon flagging the data
# as either Weekday or Weekend, while also making sure that the date
# column is formatted as the "Date" class
 
ActMon$date <- as.Date(ActMon$date, "%Y-%m-%d")
 
ActMon$weekendType <- factor(ifelse(weekdays(ActMon$date) %in% 
                                  c("Saturday", "Sunday"),
                                "Weekend", "Weekday"))
 
# Compute the mean of all steps based on the interval in which they occurred,
# then save the output as an object
T_Steps3 <- aggregate(steps ~ date, ActMon, sum)
print(summary(T_Steps3$steps))
```
 
```{r}
echo = TRUE
# Compute the mean of all steps based on the interval in which they occurred,
# then save the output as an object
Int_Act3 <- aggregate(steps ~ interval + weekendType, ActMon, mean)
 
# Plot the Weekday vs Weekend graphs of activity data
qplot(interval, steps, data = Int_Act3, facets = .~weekendType,
      geom = "line") +
  xlab("5-Minute Interval over 24 hour Period") +
  ylab("Number of Steps") +
  ggtitle("Weekday vs. Weekend")
```